---
title: "Test API connection to FAIRDOMHub"
author:
 - name: A. Blejec
   email: ablejec@nib.si
   affiliation: National Institute of Biology, Ljubljana, Slovenia
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{FAIRDOMHub-and-R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
A former version of this document is available at Available at [RPubs](http://rpubs.com/ablejec/396238)

# Introduction

I will try to connect to the FAIRDOM SEEK with the collection of their APIs.

The API description with examples is available [here](https://app.swaggerhub.com/apis/FAIRDOM/SEEK/0.1#/people/updatePerson).

It is also available in my Postman.

My final goal is to enable upload/download from pISA-tree to FAIRDOMHub. This feature will be a part of the R package `pISAR` that will enable use of pISA structure and metafile information for reproducible statistical analyses.

SEEK testing site is [here](https://testing.sysmo-db.org)

Main FAIRdomHub URL and test site urls:
```{r}
mainurl <- "https://www.fairdomhub.org/"
testurl <- "https://testing.sysmo-db.org"
```

My personal ids
```{r}
my_main_id <- 808
my_test_id <- 368
```


I will use the testing site for the development and experiments:

```{r}
siteid <- 1 # real
siteid <- 2 # test
(baseurl <- c(mainurl, testurl)[siteid])
(myid <- c(my_main_id, my_test_id)[siteid])
url(baseurl)
```

My test project id and pISA ids
```{r}
pid <- 54
iid <- 76
sid <- 82
aid <- 374
purl <- file.path(baseurl, "projects", pid)
iurl <- file.path(baseurl, "investigations", iid)
surl <- file.path(baseurl, "Studies", sid)
aurl <- file.path(baseurl, "assays", aid)
purl
# "https://testing.sysmo-db.org/projects/54"
url(purl)
url(iurl)
url(surl)
url(aurl)
```

 Check the existence

First check if the base URL exists. For this package `RCurl` can be used.

```{r}
library(RCurl)
```


```{r}
(FAIRDOMExists <-  url.exists(baseurl))
```
# Package `httr`

The use is described in [LockeData blog](https://itsalocke.com/blog/r-quick-tip-microsoft-cognitive-services-text-analytics-api/).

This solution uses packages `httr` and `jsonlite` and is based on `libcurl` (the same is used by RCurl). Package `httr` has useful vignettes that were a good guide for the example below.

```{r}
library(httr)
library(jsonlite)
```

## Get personal information

```{r}

r <- GET(file.path(baseurl,"people",myid),
         add_headers(Accept="application/json"))
names(r)
r
r$content
```
The response is structured and can be investigated with the helper functions. The response has three important parts: status, headers, and body.

```{r}
# GET ("https://testing.sysmo-db.org/people/368")
GET(paste0(baseurl,"/people/",myid))
```


### Status

Check status

```{r}
http_status(r)
```
or just a status code

```{r}
r$status_code
```

### Body

Body can be accessed with the function `content`

```{r}
cont <- content(r,"text")
cont
length(cont)
```

This gives us a text version of the JSON object. We can use built-in parsers to get other forms of data. We can parse the content into the R list object

```{r}
cont <- content(r,"parsed",type="application/json")
cont
```
### The structure of the parsed response

```{r}
names(cont)
cont$jsonapi
names(cont$data)
```

Response elements

```{r}
cont$data$id
```

List the response content

```{r}
d <- cont$data
for (nm in names(d)) {
  cat("\n",nm,"--------------\n\n")
  print(d[[nm]])
}
```

Person data are in the *attributes* part

Here is the first name of the person with id `r myid`

```{r}
cont$data$attributes$first_name
```

# Conclusion

The rudimetary R read connection with FAIRDOMHub can be established using the package `httr`. From here it should be fairly possible to write the wrapper functions for easier usage.

## Wrapper functions for `GET` and `content`

```{r}
# this function is equivaalent to a call:
# curl -X GET "https://www.fairdomhub.org/what/id" -H  "accept: application/json"
#
myGET <- function(what, id,
                   uri=baseurl, ... ){
#                  uri="https://www.fairdomhub.org", ... ){
  if(!missing(what)) uri <- paste0(uri,"/",what)
  if(!missing(id)) uri <- paste0(uri,"/",id)
  r <- GET(uri,
         add_headers(Accept="application/json"))
  cat("Status code:",r$status_code,"\n")
  invisible(r)
}

myData <- function(r,
                  part="attributes",
                  type="application/json",  ...){
  invisible(content(r,"parsed",type=type)$data[[part]])
}
myDatas <- function(r,part,
                  type="application/json",  ...){
  if(missing(part)) invisible(content(r,"parsed",type=type)$data) else
  invisible(content(r,"parsed",type=type)$data[part])
}

```


## Use of wrapper functions

Get data for person `r myid`

```{r}
myid
# https://www.fairdomhub.org/people/808
r <- myGET("people",myid)
r$status_code
```

Get the person `r myid` last name and tools

```{r}
myid
myData(r)$last_name
myData(r)$tools
```

Get person `r myid` institution id

```{r}
myData(r,"relationships")$institutions
instID <- myData(r, "relationships")$institutions$data[[1]]$id
instID
```

Get institution `\r instID` name. The details are in the *institutions* part.

```{r}
# https://www.fairdomhub.org/institutions/189
rinst <- myGET("institutions",instID)
myData(rinst)$title
```

Full institution description

```{r}
catln <- function(...) cat(...,"\n")
d <- sapply(myData(rinst),catln)
```


### Search for a person

```{r}
myname <- "Andrej Blejec"
r <- myGET("people")
d <- myDatas(r)
length(d)
names(d[[1]])
titles <- sapply(d,function(x) x$attributes$title)
head(titles)
d[[pmatch(myname,titles)]]
id <- d[[pmatch(myname,titles)]]$id
id
```

### Searh for a project

```{r}
myproject <- "testUpload"
r <- myGET("projects")
d <- myDatas(r)
length(d)
names(d[[1]])
titles <- sapply(d,function(x) x$attributes$title)
head(titles)
loc <- pmatch(myproject,titles)
d[[loc]]
d[[loc]]$attributes$title
id <- d[[loc]]$id
pid <- id
pid
```

### Searh for a investigation

```{r}
myinvestigation <- "Test1"
r <- myGET("investigations")
d <- myDatas(r)
length(d)
names(d[[1]])
titles <- sapply(d,function(x) x$attributes$title)
head(titles)
loc <- pmatch(myinvestigation,titles)
d[[loc]]
d[[loc]]$attributes$title
id <- d[[loc]]$id
iid <- id
iid
```

### GET an investigation

```{r}
extract <- function(x, what="relationships") lapply(x[[what]]
                    , FUN=function(x) if(length(x$data)>0) {
                      myDatas(myGET(x$data[[1]]$type, x$data[[1]]$id),"attributes")
                      } else NA )
#extract <- function(x, what="relationships") lapply(x[[what]]
#                    , FUN=function(x) length(x$data))

```


```{r}
id
r <- myGET("investigations", id)
d <- myDatas(r)
length(d)
names(d)
d$attributes$title
d$attributes$description
d$relationships
a <- extract(d)
str(a)
names(a)
for(nm in names(a)){
cat("\n",nm,"\n")
x <- d$relationships[[nm]]
if(length(d$relationships[[nm]]$data)>0){
a <- myGET(x$data[[1]]$type, x$data[[1]]$id)
ad <- myDatas(a)
print(ad$attributes$title)
}
}

```

List project data

```{r}
pid
r <- myGET("projects",pid)
d <- myDatas(r)
length(d)
names(d)
d$attributes$title
d$attributes$description
d$relationships
a <- extract(d)
str(a)
names(a)
for(nm in names(a)){
  cat("\n",nm,"\n")
  x <- d$relationships[[nm]]
  if(length(d$relationships[[nm]]$data)>0){
    for(j in 1:length(x$data)){
      a <- myGET(x$data[[j]]$type, x$data[[j]]$id)
      ad <- myDatas(a)
      print(ad$attributes$title)
    }
  }
}

```


## PUT

We need to create JSON objects

### Create JSON object

#### Investigation

```
{
  "data": {
    "type": "investigations",
    "attributes": {
      "title": "A Maximal Investigation",
      "policy": {
        "access": "download",
        "permissions": [
          {
            "resource": {
              "id": "809",
              "type": "projects"
            },
            "access": "manage"
          },
          {
            "resource": {
              "id": "859",
              "type": "projects"
            },
            "access": "manage"
          }
        ]
      },
      "description": "This is a more complex investigation",
      "other_creators": "Max Blumenthal, Ed Snowden"
    },
    "relationships": {
      "projects": {
        "data": [
          {
            "id": "809",
            "type": "projects"
          },
          {
            "id": "859",
            "type": "projects"
          }
        ]
      },
      "publications": {
        "data": [
          {
            "id": "23",
            "type": "publications"
          }
        ]
      },
      "creators": {
        "data": [
          {
            "id": "555",
            "type": "people"
          }
        ]
      }
    }
  }
}

```

```{r}
ir <- fromJSON('{
  "data": {
    "type": "investigations",
    "attributes": {
      "title": "A Maximal Investigation",
      "policy": {
        "access": "download",
        "permissions": [
          {
            "resource": {
              "id": "809",
              "type": "projects"
            },
            "access": "manage"
          },
          {
            "resource": {
              "id": "859",
              "type": "projects"
            },
            "access": "manage"
          }
        ]
      },
      "description": "This is a more complex investigation",
      "other_creators": "Max Blumenthal, Ed Snowden"
    },
    "relationships": {
      "projects": {
        "data": [
          {
            "id": "809",
            "type": "projects"
          },
          {
            "id": "859",
            "type": "projects"
          }
        ]
      },
      "publications": {
        "data": [
          {
            "id": "23",
            "type": "publications"
          }
        ]
      },
      "creators": {
        "data": [
          {
            "id": "555",
            "type": "people"
          }
        ]
      }
    }
  }
}')
str(ir)
ij <- toJSON(ir)
ij
```

```{r, eval=FALSE}
I <- list(
  data=list(
    type="investigations"
  , attributes=list(
      title=paste("pISA code: Test upload from R - ", Sys.time())
    , description="Uploaded from R"
  )
  , relationships=list(
      projects=list(data=data.frame(id=pid, type="projects"))
   ,  creators=list(data=data.frame(id=myid, type="people"))
  )
  )
)
I
I <- toJSON(I)
#I <- gsub("\\[","",I)
#I <- gsub("\\]","",I)
I <- gsub('("type":)\\[(".*")\\],"at','\\1\\2,"at', I)
I <- gsub('("title":)\\[(".*")\\],"de','\\1\\2,"de', I)
I <- gsub('("description":)\\[(".*")\\]},','\\1\\2},', I)
I
```
Or nicer:

```{r}
prettify(I)
```

Escape quotes - not a remedy, not needed
```{r, eval =FALSE}
II <- gsub('\\"','\\\\"',I)
II
```


## Create new investigaton (POST)


```{r}
uri <- modify_url(baseurl,path="investigations")
uri
response <- POST(uri
    , authenticate("ablejec","testni.1234",type = "basic")
    , body = I
    , encode="json"
    , accept("application/json")
    , content_type_json()
    )
res <- fromJSON(content(response, as = "text"))
str(res)
```

Open in the web browser
```{r}
if(interactive()) shell.exec(file.path(res$data$meta$base_url,res$data$links$self))
```



```
system('curl -X POST "https://testing.sysmo-db.org/investigations" -H  "accept: application/json" -H  "Content-Type: application/json" -u "ablejec:testni.1234" -d "{\"data\":{\"type\":\"investigations\",\"attributes\":{\"title\":\"Title 0.116822839947417\",\"description\":\"Description\"}},\"relationships\":{\"projects\":{\"id\":\"54\",\"type\":\"projects\"},\"creators\":{\"id\":368,\"type\":\"people\"}}}"')
```

```
system('curl -X POST "https://testing.sysmo-db.org/investigations" -H  "accept: application/json" -H  "Content-Type: application/json" -u "ablejec:testni.1234" -d "{\"data\":{\"type\":\"investigations\",\"attributes\":{\"title\":\"Title 0.116822839947417\",\"description\":\"Description\"}},\"relationships\":{\"projects\":{\"id\":\"54\",\"type\":\"projects\"},\"creators\":{\"id\":368,\"type\":\"people\"}}}"')
```



From RCurl

```{r}
r <- myGET("investigations")
d <- myDatas(r)
length(d)
names(d)
d$attributes$title
d$attributes$description
d$relationships
a <- extract(d)
str(a)
names(a)
for(nm in names(a)){
cat("\n",nm,"\n")
x <- d$relationships[[nm]]
if(length(d$relationships[[nm]]$data)>0){
a <- myGET(x$data[[1]]$type, x$data[[1]]$id)
ad <- myDatas(a)
print(ad$attributes$title)
}
}
```
This works in Postman
```
{
  "data": {
    "type": "investigations",
    "attributes": {
      "title": "A Maximal Investigation",
      "policy": {
        "access": "download",
        "permissions": [
          {
            "resource": {
              "id": "54",
              "type": "projects"
            },
            "access": "manage"
          }
        ]
      },
      "description": "This is a more complex investigation",
      "other_creators": "Max Blumenthal, Ed Snowden"
    },
    "relationships": {
      "projects": {
        "data": [
          {
            "id": "54",
            "type": "projects"
          }
        ]
      },
        "creators": {
        "data": [
          {
            "id": "368",
            "type": "people"
          }
        ]
      }
    }
  }
}
```

Tudi moj primer

```
{
    "data": {
        "type": "investigations",
        "attributes": {
            "title": "Title 0.4220704715696",
            "description": "Description"
        },
        "relationships": {
        "projects": {
            "data": [
                {
                    "id": "54",
                    "type": "projects"
                }
            ]
        },
        "creators": {
            "data": [
                {
                    "id": "368",
                    "type": "people"
                }
            ]
        }
    }
    }

}
```

List project data

```{r}
pid
r <- myGET("projects",pid)
d <- myDatas(r)
length(d)
names(d)
d$attributes$title
d$attributes$description
d$relationships
a <- extract(d)
str(a)
names(a)
for(nm in names(a)){
  cat("\n",nm,"\n")
  x <- d$relationships[[nm]]
  if(length(d$relationships[[nm]]$data)>0){
    for(j in 1:length(x$data)){
      a <- myGET(x$data[[j]]$type, x$data[[j]]$id)
      ad <- myDatas(a)
      print(ad$attributes$title)
    }
  }
}

```

## CReate and post documents

```{r}
document_skeleton <- '{
  "data": {
    "type": "documents",
    "attributes": {
      "title": "A Maximal Document",
      "description": "This is the description",
      "tags": [
        "tag1",
        "tag2"
      ],
      "license": "CC-BY-4.0",
      "other_creators": "John Smith, Jane Smith",
      "content_blobs": [
        {
          "original_filename": "a_pdf_file.pdf",
          "content_type": "application/pdf"
        }
      ],
      "policy": {
        "access": "download",
        "permissions": [
          {
            "resource": {
              "id": "359",
              "type": "projects"
            },
            "access": "edit"
          }
        ]
      }
    },
    "relationships": {
      "creators": {
        "data": [
          {
            "id": "234",
            "type": "people"
          }
        ]
      },
      "projects": {
        "data": [
          {
            "id": "359",
            "type": "projects"
          }
        ]
      },
      "assays": {
        "data": [
          {
            "id": "38",
            "type": "assays"
          }
        ]
      }
    }
  }
}'
document_skeleton
II <- fromJSON(document_skeleton, simplifyVector = TRUE)
II
```


```{r}
content_type <- "text/plain"
apath <- "D:/DEJAVNOSTI/OMIKE/pISA-projects/_p_biopharm.si/_I_Task1/_S_doktoratiUM/_A_080-LDA-Ref-R"
original_filename <- "_ASSAY_METADATA.TXT"
document_title <- paste("Assay metadata:",Sys.time())
document_description <- "Details about assay"
#
II$data$attributes$title <- document_title
II$data$attributes$description <- document_description
II$data$relationships$creators$data$id <- myid
II$data$relationships$creators$data
II$data$relationships$projects$data$id <- pid
II$data$relationships$projects$data
II$data$attributes$policy$permissions$resource$id <- pid
II$data$attributes$policy$permissions
II$data$attributes$content_blobs$content_type <- content_type
II$data$attributes$content_blobs$original_filename <- original_filename
II$data$attributes$content_blobs
II$data$attributes$other_creators <- ""
II$data$attributes$tags <- c("pISA", "metadata")
II$data$relationships$assays$data$id <- aid
II
I <- toJSON(II,auto_unbox = TRUE)
I
```

## Create new document (POST)


```{r}
uri <- modify_url(baseurl,path="documents")
uri
response <- POST(uri
    , authenticate("ablejec","testni.1234",type = "basic")
    , body = I
    , encode="json"
    , accept("application/json")
    , content_type_json()
    )
res <- fromJSON(content(response, as = "text"))
str(res)
res$data$id
original_filename <- res$data$attributes$content_blobs$original_filename
blob_link <- res$data$attributes$content_blobs$link
blob_link
```

Open in the web browser
```{r}
if(interactive()) shell.exec(file.path(res$data$meta$base_url,res$data$links$self))
```


### Fill the content blob with data

```{r}

fpath <- file.path(apath, original_filename)
file_content <- paste(readLines(fpath),collapse="\n")
cat(file_content)
```

```{r}
uri <- blob_link
uri
response <- PUT(uri
    , authenticate("ablejec","testni.1234",type = "basic")
    , body = upload_file(fpath)
    , encode="json"
    , accept("*.*")
    )
res <- fromJSON(content(response, as = "text"))
str(res)
```

Open in the web browser
```{r}
if(interactive()) shell.exec(blob_link)
```



# Links

[Best practices for API packages](https://cran.r-project.org/web/packages/httr/vignettes/api-packages.html)

