---
title: "Upload to FAIRDOMHub with seekr"
author: "A. Blejec"
date: "11. 10. 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
.testing <- TRUE
```

## Introduction

I will organize a batch upload to FAIRDOMHub. For the operation I have prepared the R package `seekr`, available at [GitHub](https://github.com/NIB-SI/seekr). Package operations are based on SEEK API [documentation]{https://fairdomhub.org/api}.

```{r}
library(pisar)
library(rio)
library(jsonlite)
library(httr)
library(seekr)
```


## Prepare pISA environment

Package `seekr` contains demo pISA-tree project. For this demonstration we can use the demo project. It is assumed, that our working directory is a layer of the pISA project, in our case we will start in an assay.

```{r}
astring <- "_p_Demo/_I_Test/_S_Show/_A_Work-R/"
oldwd <- setwd(system.file("extdata", astring, package="seekr"))
oldwd
dir()
```


Initialize the pISA-tree. Function `pisa` will complain, if the working directory is not a pISA layer.

```{r}
pini <- pisa()
```

Let us check some of the infos gathered:

Assay name:
```{r}
.aname
# or
pini$A$name
```

Study information

```{r}
pini$S
```

Study and project descriptions
```{r}
getMeta(.smeta,"Description")
getMeta(.pmeta,"Description")
```


## Prepare study files

SEEK supports only files in assays. Study files need to be copied to an assay.


```{r}
snames <- dir(.iroot, pattern="^_S_"
              , full.names = FALSE
              , recursive=FALSE
              , include.dirs=TRUE)
snames
#snames <- snames[grepl("^_S_",snames)]
for(sname in snames) {
  asname <- paste0("_A_",sname,"-files")
  aspath <- file.path(.iroot,sname,asname)
  if(!dir.exists(aspath)) dir.create(aspath)
  #file.copy(file.path(.iroot,sname,"*.*"), aspath)
}
```

Get the directory listing of all studies in the parent investigation
and exclude files listed in file `seekignore.txt`. In our case, seekignore file is in the investigation layer.

```{r}
# Get SEEK ignore file
ignore_file <- file.path(.iroot,"seekignore.txt")
if(file.exists(ignore_file))
seekignore <- read.table(ignore_file)[,1] else seekignore <- ""
seekignore

# Create directory listing

dirs <-  dir(.iroot
             , full.names = FALSE
             , recursive=TRUE
             , include.dirs=FALSE)
head(dirs,10)
length(dirs)

# Ignore files with patterns in seekignore file
dirlength <- length(dirs)
for(pat in toupper(seekignore)){
  ldir <- length(dirs)
  at <- gsub("\\.","\\\\.",pat)
  dirs <- dirs[!grepl(pat,toupper(dirs))]
  cat("Ignoring" , ldir-length(dirs), "files [",pat," ]\n")
}
cat("Files to upload:", length(dirs), "/",dirlength)

# Keep only files in Assays

head(dirs)

dirs <- dirs[grepl("/_A_", dirs)]

head(dirs)
tail(dirs)
cat("Files to upload:", length(dirs), "/",dirlength,"\n")
```

```{r eval=FALSE, include=FALSE, echo=FALSE}
dirs=c("xx/other/","bla/other-file.txt")
pat="other/"
dirs[!grepl(toupper(pat),toupper(dirs),fixed=TRUE)]
grep(toupper(pat),toupper(dirs),fixed=TRUE, value=TRUE, invert=TRUE)
glob2rx(pat,trim.head=TRUE,trim.tail=TRUE)
```

## Prepare connection with FAIRDOMHub

Restart the log file.

```{r}
skLog("Starting upload for demo project", append=FALSE)
```

Initialize connection with SEEK cloud service. Cloud service url and project creator name and SEEK service id are expected in the project metadata file. They should be stored in options().

```{r}
options(.sk$test)
# options(.sk$main)
skOptions()
```


Here I am showing a part of the settings (to not reveal my password).

Find creator's FAIRDOMHub id. Project creator name should be spelled exactly as in the FAIRDOMHub account.

```{r}
meta <- pini$p$meta
crid <- skFindId("people",meta[grepl("creator",meta$Key),]$Value)["id"]
crid
crid <- skFindId("people","Andrej Blejec")["id"]
crid
if(crid == 0 ) crid <- 368
myid <- crid
myid
```


```{r}
options(sk.myid = crid)

```

One might be interested in the institution the person belongs. SEEK API returns a JSON structure, which is converted into an R list. One can traverse the list in usual way to get to the relevant data.

```{r}
resp <- skContent( skRead("people",  getOption( "sk.myid" ) ) )$relationships$institutions$data[[1]]
resp
str(resp)
inst <- skRead(resp$type,resp$id)
inst
skContent(inst)$attributes$title
```


### Get project and Investigation ids:

Usually, the project on FAIRDOMHub will be created interactively by the programme admin. First, we will check the existence of our demo project `r pini$p$name`on the FAIRDOMHub.

```{r}
pini$p$name
pinfo <- skFindId("projects", pini$p$name)
pinfo
# project identification number
pid <- pinfo["id"]
# project name
pname <-  pinfo["title"]
pname
pini$p$name == pname
#
```
The local project name and the name on the cloud match. On the FAIRDOMHub, project components have identification numbers. Our project has the identification number `r pid`:

```{r}
(pid <- pinfo["id"])
```
Project id is part of the path to the project on the cloud:

```{r}
skRead("projects",pid)$url
```





```{r}
iinfo <- skFindId("investigations", pini$I$name)
#iinfo <- c(id=325,type="investigation",title=pini$I$name)
iinfo
iid <- iinfo["id"]
iname <- iinfo["title"]
# project id (todo: make a function)
prdata <- skRead("projects", pid)
prid <- prdata$content$relationships$programmes$data[[1]][["id"]]
prid
```

If the layer (investigation) does not exist, the function skFindId returns zero in the place of 'id'.

If it is so, one can create the appropriate layer.

```{r eval=(iid == 0)}
skCreate("investigations", meta=.imeta)
skFindId("investigations", pini$I$name)
```

Update seekr options.

```{r}
options(prid)
options(pid)
options(iid)
skOptions()
```

Parse file path into study, assay, filename

## Pick the files to upload

```{r}
uplimit <- 3*10^9
#ind <- 205  # random line numbers
#(ind <- which(grepl("_S_03_stCuSTr/_A_01_evigene",dirs,ignore.case=TRUE)))
ind <- 1:length(dirs)
nind <- length(ind)
# For testing
.testing <- FALSE
if(.testing){
ind <- sample(ind,10)
nind <- length(ind)
dirs <- dirs[ind]
ind <- 1:length(dirs)
}
```

Loop through list of files and upload

```{r}
for(i in ind){
  i <- 1
  x <- dirs[i]
 cat("\n\n|", rep("-",50*i/nind-1),">", rep(" ",50*(nind-i)/nind),"| ", i, "/", nind, "\n", x ,"\n\n", rep("-",50),"\n", sep="")
sname <- gsub("^.*(_S_[^/]*)/_A.*","\\1",x)
aname <- gsub("^.*/(_A_[^/]*)/.*","\\1",x)
fname <- gsub("^.*/(_[AS]_[^/]*)/(.*)","\\2",x)
.sname <- sname
.aname <- aname
.aname
#
# Check if the study exists, create if not
#
# Find Study id
sid <- skFindId("studies",sname)["id"]
sid
sresp <- skRead("studies",sid)

if(sresp$content$relationships$investigation$data$id!=iid) sid <- 0
skIni(pid=pid,iid=iid)
smeta <- readMeta(file.path(.iroot,sname))
meta <- list(Title = sname , Description = pasteMeta(smeta) )
#cat(meta$Description)
if(sid==0) { sresp <- skCreate("studies",meta)
  print(sresp)
  sid <- sresp$id
  }
sid
#skRead("studies",sid)
skIni(pid=pid,iid=iid,sid=sid)
#
# Check if the assay exists, create if not
#
aid <- skFindId("assays",aname)["id"]
aresp <- skRead("assays",aid)
aid
#if(aresp$content$relationships$study$data$id!=sid) aid <- 0
#if(aresp==404) aid <- 0
aid
ameta <- readMeta(file.path(.iroot,sname,aname))
if(getMeta(ameta,"Assay Class")if(FALSE) {
meta <- list(Title = aname
             , Description = paste(
               getMeta(meta, "Title")
               ,getMeta(meta,"Description")
               , sep=" \n "
               )
)}
meta <- list(Title = aname , Description = pasteMeta(ameta) )
meta
if(aid==0) { aresp <- skCreate("assays",meta, class=skClass)
print(aresp)
aid <- aresp$id
}
aid
skIni(pid=pid,iid=iid,sid=sid,aid=aid)
#
# Upload file
#
options("seekr")
if(.testing) {
  (skRead("studies", options()$seekr$sid))
  (skRead("assays" , options()$seekr$aid))
}
# create and upload file
x
fname
meta <- list(
       Title=paste0("/",fname)
     , Description=paste(pini$p$name,pini$I$name,sname,aname,sep="/")
     )
if(toupper(fileType(fname))    regexpr("METADATA",fname)>0) {
  type <- "documents" } else {type <- "data_files"}
s <- skCreate(type, meta, file= fname)
if(.testing) print(s)
#
f_size <- file.size(file.path(.iroot,x))
if(!is.na(f_size)){
if(f_size< uplimit) {
cat("Uploading ", fname)
f <- skUpload(s, file.path(.iroot,x))
} else { f <- paste( "Not uploaded, File too large", f_size ,">",  uplimit/10^9,"GB") } } else  { f <-  x
        cat("Length NA, not uploaded: ")
}
print(f)
}
```







