---
title: "HowTo Use seekr"
author: "A. Blejec"
date: "Last compiled on `r format(Sys.time(), '%Y-%m-%d')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
.testing <- TRUE
```

## Introduction

I will organize a batch upload to FAIRDOMHub. For the operation I have prepared the R package `seekr`, available at [GitHub](https://github.com/NIB-SI/seekr). Package operations are based on SEEK API [documentation]{https://fairdomhub.org/api}.

```{r}
library(pisar)
#library(rio)
library(jsonlite)
library(httr)
library(seekr)
```


## Prepare pISA environment

Package `seekr` contains demo pISA-tree project. For this demonstration we can use the demo project. It is assumed, that our working directory is a layer of the pISA project, in our case we will start in an assay.

```{r}
astring <- "_p_Demo/_I_Test/_S_Show/_A_Work-R/"
oldwd <- setwd(system.file("extdata", astring, package="seekr"))
oldwd
dir()
```


Initialize the pISA-tree. Function `pisa` will complain, if the working directory is not a pISA layer.

```{r}
pini <- pisa()
```

Let us check some of the infos gathered:

Assay name:
```{r}
.aname
# or
pini$A$name
```

Study information

```{r}
pini$S
```

Study and project descriptions
```{r}
getMeta(.smeta,"Description")
getMeta(.pmeta,"Description")
```


## Prepare connection with FAIRDOMHub

Restart the log file.

```{r}
skLog("Starting upload for demo project", append=!FALSE)
```

Initialize connection with SEEK cloud service. Cloud service url and project creator name and SEEK service id are expected in the project metadata file. They should be stored in options().

```{r}
options(.sk$test)
# options(.sk$main)
skOptions()
```


Here I am showing a part of the settings (to not reveal my password).

Find creator's FAIRDOMHub id. Project creator name should be spelled exactly as in the FAIRDOMHub account.

```{r}
meta <- pini$p$meta
crid <- skFindId("people",meta[grepl("creator",meta$Key),]$Value)["id"]
crid
crid <- skFindId("people","Andrej Blejec")["id"]
crid
if(crid == 0 ) crid <- 368
myid <- crid
myid
```

Store creator id in options.

```{r}
options(sk.myid = crid)

```

One might be interested in the institution the person belongs. SEEK API returns a JSON structure, which is converted into an R list. One can traverse the list in usual way to get to the relevant data.

```{r}
resp <- skContent( skRead("people",  getOption( "sk.myid" ) ) )$relationships$institutions$data[[1]]
resp
str(resp)
inst <- skRead(resp$type,resp$id)
inst
skContent(inst)$attributes$title
```


### Get programme, project and Investigation ids:

On FAIRDOMHub projects are parts of programmes. Find the id of programme
connected to pISA-tree:

```{r}
skFindId("prog","pISA-tree")
```

Usually, the project on FAIRDOMHub will be created interactively by the programme admin. First, we will check the existence of our demo project `r pini$p$name`on the FAIRDOMHub.

```{r}
pini$p$name
pinfo <- skFindId("projects", pini$p$name)
pinfo
# project identification number
pid <- pinfo["id"]
# project name
pname <-  pinfo["title"]
pname
pini$p$name == pname
#
```
The local project name and the name on the cloud match. On the FAIRDOMHub, project components have identification numbers. Our project has the identification number `r pid`:

```{r}
(pid <- pinfo["id"])
```
Project id is part of the path to the project on the cloud:

```{r}
skRead("projects",pid)$url
```

Find investigation id

```{r}
iinfo <- skFindId("investigations", pini$I$name)
#iinfo <- c(id=325,type="investigation",title=pini$I$name)
iinfo
iid <- iinfo["id"]
iname <- iinfo["title"]
```

Each object/layer has links/ids to associated objects. The programme id can be found in the result of skRead (read object):

```{r}
# project id (todo: make a function)
prdata <- skRead("projects", pid)
prid <- prdata$content$relationships$programmes$data[[1]][["id"]]
prid
```

More elegant way might be a call to skRelationships.

```{r}
skRelated("proj",pid,"program")
```

The first line lists the investigated object info and then list of zero or more related objects of the type specified by the third argument.

If the layer (investigation) does not exist, the function skFindId returns zero in the place of 'id'. There is a convenience function \fct{skExists} that checks existence of a layer of object.

```{r}
iid
pini$I$name
skExists("inv",pini$I$name,  "proj", verbose=TRUE)
```
If it is so, one can create the appropriate layer. Created object id is stored in options or can be found by a call to function \fct{skFindId}.

```{r eval=(iid == 0)}
if( !skExists("inv", pini$I$name , "proj" )) 
     skCreate("investigations", meta=.imeta)
skFindId("investigations", pini$I$name)
options("sk.iid")
skExists("inv", pini$I$name,  "proj", verbose=TRUE)
skRelated("inv", select="proj", verbose=TRUE)
```


So far,  in SEEK files can be associated only to assays. As a workaround, metadata are recorded in the description field of attributes.. This is how the metadata can be retrieved:

```{r}
r <- skRead("investigations", getOption("sk.iid"))
r <- skContent(r)
names(r$attributes)
(desc <- r$attributes$description)
```

Metadata are a string with tabs and new-lines, so it can be typed to the console.

```{r}
cat(desc)
```

One way to convert metadata string to a data frame:

```{r}
cat(desc, file="clipboard")
im <- read.table("clipboard", sep="\t", col.names=c("Key","Value"))
im
```

Package pisar has a method to nicely print metadata.

```{r}
 class(im)=c("pISAmeta","Dlist","data.frame")
 im
```

The last touched object id is stored in the appropriate seekr options.

```{r}
print.simple.list(skOptions())
```

If you wants to be sure, you can update the seekr options.

```{r}
options("sk.prid")
options("sk.pid")
options("sk.iid")
skOptions()
```

# Upload

## Prepare study files

SEEK supports only files in assays. Study files need to be copied to an assay.


```{r}
snames <- dir(.iroot, pattern="^_S_"
              , full.names = FALSE
              , recursive=FALSE
              , include.dirs=TRUE)
snames
#snames <- snames[grepl("^_S_",snames)]
for(sname in snames) {
  asname <- paste0("_A_",sname,"-files")
  aspath <- file.path(.iroot,sname,asname)
  if(!dir.exists(aspath)) dir.create(aspath)
  #file.copy(file.path(.iroot,sname,"*.*"), aspath)
}
```

###!Get the directory listing of all studies in the parent investigation
###!and exclude files listed in file `seekignore.txt`. In our case, seekignore file is in the investigation layer.
###!
###!```{r}
###!# Get SEEK ignore file
###!ignore_file <- file.path(.iroot,"seekignore.txt")
###!if(file.exists(ignore_file))
###!seekignore <- read.table(ignore_file)[,1] else seekignore <- ""
###!#seekignore <- c("_A_Work-R")
###!seekignore
###!
###!# Create directory listing
###!
###!dirs <-  dir(.iroot
###!             , full.names = FALSE
###!             , recursive=TRUE
###!             , include.dirs=FALSE)
###!head(dirs,10)
###!length(dirs)
###!
###!# Ignore files with patterns in seekignore file
###!dirlength <- length(dirs)
###!for(pat in toupper(seekignore)){
###!  ldir <- length(dirs)
###!  at <- gsub("\\.","\\\\.",pat)
###!  dirs <- dirs[!grepl(pat,toupper(dirs))]
###!  cat("Ignoring" , ldir-length(dirs), "files [",pat," ]\n")
###!}
###!cat("Files to upload:", length(dirs), "/",dirlength)
###!
###!# Keep only files in Assays
###!
###!head(dirs)
###!
###!dirs <- dirs[grepl("/_A_", dirs)]
###!
###!head(dirs)
###!tail(dirs)
###!dirs
###!cat("Files to upload:", length(dirs), "/",dirlength,"\n")
###!```
###!
###!```{r eval=FALSE, include=FALSE, echo=FALSE}
###!if(FALSE){
###!dirs=c("xx/other/","bla/other-file.txt")
###!pat="other/"
###!dirs[!grepl(toupper(pat),toupper(dirs),fixed=TRUE)]
###!grep(toupper(pat),toupper(dirs),fixed=TRUE, value=TRUE, invert=TRUE)
###!glob2rx(pat,trim.head=TRUE,trim.tail=TRUE)
###!}
###!```

## Prepare list of files

The most important task is to prepare the list of files to be uploaded. Typically one will upload files in one investigation, usually with all related branches (studies and assays). The metadata key "Upload to FAIRDOMHub" (Yes/No) controls if the layer (together with all subordinated layers) should be included or ignored by the upload process. In addition,  specific files or paths can be declared as ignored by the upload process.
File `seekignore.txt` in the investigation layer contains file names, directory names or name patterns that are used to filter and exclude the list of files to be uploaded. An example of the seekignore file is here:

```{r}
seekignore <- readLines(file.path(.iroot, "seekignore.txt"))
seekignore
```

Lines three and four have slash (/) at the end to ignore directories. Line five and six will prevent upload of batch files and files with specific name (in any directory). One can add other patterns into the seekignore.txt file before running the code below.

Function `skFilesToUpload` will first create a list of files in root directory. Then, it will read the seekignore.txt and extend it with patterns of layers that should not be uploaded according to metadata key "Upload to FAIRDOMHub".
In addition, seekignre list will be appended with the list of empty README.MD files that might clutter the FAIRDOMHub list of files.
Then it will filter it according to declarations in the combined seekignore pattern list.

```{r}
forUpload <- skFilesToUpload( root = pini$I$root)
```

Final list of files for upload is

```{r}
head(forUpload)
tail(forUpload)
dirs <- forUpload
```



Parse file path into study, assay, filename

## Pick the files to upload

```{r}
uplimit <- 3*10^9
#ind <- 205  # random line numbers
#(ind <- which(grepl("_S_03_stCuSTr/_A_01_evigene",dirs,ignore.case=TRUE)))
ind <- 1:length(dirs)
nind <- length(ind)
# For testing
.testing <- FALSE
if(.testing){
ind <- sample(ind,10)
nind <- length(ind)
dirs <- dirs[ind]
ind <- 1:length(dirs)
}
```

Loop through list of files and upload

```{r}
ind <- 1:3
for(i in ind){
  x <- dirs[i]
 cat("\n\n|", rep("-",50*i/nind-1),">", rep(" ",50*(nind-i)/nind),"| ", i, "/", nind, "\n", x ,"\n\n", rep("-",50),"\n", sep="")
sname <- gsub("^.*(_S_[^/]*)/_A.*","\\1",x)
if(sname==x) sname <- "" else sname <- sname
aname <- gsub("^.*/(_A_[^/]*)/.*","\\1",x)
if(aname==x) aname <- "" else aname <- aname
fname <- gsub("^.*/(_[AS]_[^/]*)/(.*)","\\2",x)
.sname <- sname
.aname <- aname
print(c(.sname, .aname, fname))
#
# Check if the study exists, create if not
#
if(sname!=""){
# Find Study id

sid <- skFindId("studies",sname)["id"]
sid
sresp <- skRead("studies",sid)

if(sresp$content$relationships$investigation$data$id!=iid) sid <- 0
skIni(pid=pid,iid=iid)
smeta <- readMeta(file.path(.iroot,sname))
meta <- list(Title = sname , Description = pasteMeta(smeta) )
#cat(meta$Description)
if(sid==0) { sresp <- skCreate("studies",meta)
  print(sresp)
  sid <- sresp$id
  }

sid
#skRead("studies",sid)
skIni(pid=pid,iid=iid,sid=sid)
#
# Check if the assay exists, create if not
#
aid <- skFindId("assays",aname)["id"]
aresp <- skRead("assays",aid)
aid
#if(aresp$content$relationships$study$data$id!=sid) aid <- 0
#if(aresp==404) aid <- 0
aid
ameta <- readMeta(file.path(.iroot,sname,aname))
# ToDO what are the ifs?????
#if(getMeta(ameta,"Assay Class")
if(FALSE) {
meta <- list(Title = aname
             , Description = paste(
               getMeta(meta, "Title")
               ,getMeta(meta,"Description")
               , sep=" \n "
               )
)}
meta <- list(Title = aname , Description = pasteMeta(ameta) )
meta
if(aid==0) { aresp <- skCreate("assays",meta, class=skClass)
print(aresp)
aid <- aresp$id
}
aid
  } else {
  sid <- ""
  aid <- ""
  }
  skSetOption("studies", sid)
  skSetOption("assays",aid)
#skIni(pid=pid,iid=iid,sid=sid,aid=aid)
}
print.simple.list(skOptions())
}

```
```{r}

#
# Upload file
#
options("seekr")
if(.testing) {
  (skRead("studies", options()$seekr$sid))
  (skRead("assays" , options()$seekr$aid))
}
# create and upload file
x
fname
meta <- list(
       Title=paste0("/",fname)
     , Description=paste(pini$p$name,pini$I$name,sname,aname,sep="/")
     )
## ToDo
if(toupper(fileType(fname))    regexpr("METADATA",fname)>0) {
  type <- "documents" } else {type <- "data_files"}
s <- skCreate(type, meta, file= fname)
if(.testing) print(s)
#
f_size <- file.size(file.path(.iroot,x))
if(!is.na(f_size)){
if(f_size< uplimit) {
cat("Uploading ", fname)
f <- skUpload(s, file.path(.iroot,x))
} else { f <- paste( "Not uploaded, File too large", f_size ,">",  uplimit/10^9,"GB") } } else  { f <-  x
        cat("Length NA, not uploaded: ")
}
print(f)
}
```







